Computing the Hazard Ratios associated with Explanatory Variables using Machine Learning Models of Survival Data
Authors: Sameer Sundrani and James Lu,*

Affiliations: 
1) Modeling & Simulation/Clinical Pharmacology, Genentech, South San Francisco, California 
2) Biomedical Computation, Schools of Engineering and Medicine, Stanford University, Stanford, California 
*Corresponding author. Email: lu.james@gene.com 


This R Notebook contains all the code used for CoxPH / XGBoost Comparisons for simulated survival data only.
the following sections:
(1) Set Directory
(2) Packages and Functions 
(3) Generate Simulated Data

Set working directory for saving files
```{r}
#Location of where to store resulting files and figures, everything will be saved here
dir <- paste0(dirname(rstudioapi::getActiveDocumentContext()$path),"/")
dir.create("data_sim")
dir.create("results_sim")


#global constants used for parsing dataframes
coef_val <- 7 #location in the CoxPH table for coef
ci_val <- 8 #location in the CoxPH table for CI
```

Include These Packages for Analysis
```{r}
#package needed for Breast Cancer data
library("condSURV")

#packages specific for Nature Data b/c we need to clean the data
library(rpart)
library(mice)
library(dummies)

#general packages
library(tidyverse)
library(extraFunctions)
library(haven)
library(rlang)
library(aod)
library(ggpubr)
library(caret)
library(dataPreparation)
library("ggfortify")
library("corrplot")
library("Hmisc")
library("survival")
library("survminer")
library("forestmodel")
library(coxed)
```

Function for generating tables.
```{r}
#FUNCTION for full results table (all covs)
# PARAMS:
#   coxfit: CoxPH summary
#   listCoef: List of coefficient variables
res_tab <- function(coxfit, listCoef){
  df <- as_tibble(summary(coxfit)[coef_val]$coef) %>% 
  left_join(as_tibble(summary(coxfit)[ci_val]$conf.int),  by = 'exp(coef)') %>% 
  cbind(Covariate = listCoef) %>% 
  select(Covariate, Beta = 'coef', HR = 'exp(coef)', 
         p.val = 'Pr(>|z|)', LB = 'lower .95', UB = 'upper .95') %>% 
  mutate(logLB = log(LB), logUB = log(UB))
}
```


#Simulated Survival data
```{r}
#Make random dataset
X_creation = cbind(rbinom(850, 1, 0.5), rbinom(850, 1, 0.5), rbinom(850, 1, 0.5))

#N = patients, T = range of survival times
simdata <- sim.survdata(X = X_creation, T = 10000, censor=.2 ,beta=c(1, -2,2), num.data.frames=1)
beta_known <- simdata$betas

#GENERATE dataset of binaries
simdf <- simdata$data %>% 
  mutate(
  #        X1 = ifelse(X1 >= median(X1), 1, 0), 
  #        X2 = ifelse(X2 >= median(X2), 1, 0), 
  #        X3 = ifelse(X3 >= median(X3), 1, 0),
         failed = ifelse(failed == 'TRUE', 1, 0)) %>% 
  select(coef1 = X1, coef2 = X2, coef3 = X3, time = y, status = failed) 
write_csv(simdf, paste0(dir, "data_sim/", "3coef_SIMDATA.csv"))

coxmodel <- coxph(Surv(time, status) ~ ., data=simdf)
summary(coxmodel)

results_table_sim <- res_tab(coxmodel, c("coef1", "coef2", "coef3")) %>% 
select(Covariate, HR, LB, UB) %>% 
write_csv(paste0(dir, "results_sim/", "summary_coxPH_SIMDATA.csv"))

simdf_forXGB <- simdf %>% 
  mutate(SURV_DAY_XGB = ifelse(status == 1, time, -1*time)) %>% 
  select(coef1, coef2, coef3, SURV_DAY_XGB)%>% 
write_csv(paste0(dir, "data_sim/", "xgb_SIMDATA.csv"))


#5-Fold CV
#From: https://stats.stackexchange.com/questions/61090/how-to-split-a-data-set-to-do-10-fold-cross-validation 
#Randomly shuffle the data
dataCV <-simdf
dataCV <-dataCV[sample(nrow(dataCV)),]

#Create 5 equally size folds
folds <- cut(seq(1,nrow(dataCV)),breaks=5,labels=FALSE)
listCs <- list()
#Perform 5 fold cross validation
for(i in 1:5){
    #Segement your data by fold using the which() function 
    testIndexes <- which(folds==i,arr.ind=TRUE)
    testData <- dataCV[testIndexes, ]
    trainData <- dataCV[-testIndexes, ]
    trained_cox <- (coxph(Surv(time, status) ~ ., data = trainData))
    listCs <-
      c(listCs, survConcordance(
        Surv(time, status) ~ predict(trained_cox, testData),
        testData
      )$concordance)
    
}
listCs
#Print values
for (i in 1:5){
  print((listCs[i]$concordant[1]))
}

#get 5-Fold CV results for plots
cox5fcv <-c(0.7946081, 0.7872674, 0.7634267, 0.8172029, 0.7870754)
meanSIM <- mean(cox5fcv)
sdSIM = sd(cox5fcv)
meanSIM
sdSIM
```

