Computing the Hazard Ratios associated with Explanatory Variables using Machine Learning Models of Survival Data
Authors: Sameer Sundrani and James Lu,*

Affiliations: 
1) Modeling & Simulation/Clinical Pharmacology, Genentech, South San Francisco, California 
2) Biomedical Computation, Schools of Engineering and Medicine, Stanford University, Stanford, California 
*Corresponding author. Email: lu.james@gene.com 


This R Notebook contains all the code used for CoxPH / XGBoost Comparisons. 

The code is divided into the following sections:
(1) Set Directory
(2) Packages and Functions 
(3) Breast Cancer Data Analysis 
(4) Colon Cancer Data Analysis 
(5) Pan-Cancer Data Analysis

Set working directory for saving files
```{r}
#Location of where to store resulting files and figures, everything will be saved here
dir <- paste0(dirname(rstudioapi::getActiveDocumentContext()$path),"/")
dir.create("data_process")
dir.create("data_normalized")
dir.create("figures")
dir.create("results")

#global constants used for parsing dataframes
coef_val <- 7 #location in the CoxPH table for coef
ci_val <- 8 #location in the CoxPH table for CI
```

Include These Packages for Analysis
```{r}
#package needed for Breast Cancer data
library("condSURV")

#packages specific for Nature Data b/c we need to clean the data
library(rpart)
library(mice)
library(dummies)

#general packages
library(tidyverse)
library(extraFunctions)
library(haven)
library(rlang)
library(aod)
library(ggpubr)
library(caret)
library(dataPreparation)
library("ggfortify")
library("corrplot")
library("Hmisc")
library("survival")
library("survminer")
library("forestmodel")
library(coxed)
```

Functions for generating tables, KM curves, etc.
```{r}
#FUNCTION for full results table (all covs)
# PARAMS:
#   coxfit: CoxPH summary
#   listCoef: List of coefficient variables
res_tab <- function(coxfit, listCoef){
  df <- as_tibble(summary(coxfit)[coef_val]$coef) %>% 
  left_join(as_tibble(summary(coxfit)[ci_val]$conf.int),  by = 'exp(coef)') %>% 
  cbind(Covariate = listCoef) %>% 
  select(Covariate, Beta = 'coef', HR = 'exp(coef)', 
         p.val = 'Pr(>|z|)', LB = 'lower .95', UB = 'upper .95') %>% 
  mutate(logLB = log(LB), logUB = log(UB))
}

#FUNCTION for summary table and to generate a summary table 
#of appropriate medians for NON-BINARY COVS.
# PARAMS:
#   inputNormalizedDf: normalized data
#   nonbinDf: summary table of the nonbinary covariates
#   listCoef:List of coefficent variables
nonbin_summary <- function(inputNormalizedDf, nonbinDf, listCoef){
  totalList <- list()
  i <- 1
  for (var in listCoef){
    gMean <-
      mean(as.numeric(unlist(inputNormalizedDf[inputNormalizedDf[[var]] >= median(inputNormalizedDf[[var]], na.rm = T), var])), na.rm = T)
    lMean <-
      mean(as.numeric(unlist(inputNormalizedDf[inputNormalizedDf[[var]] < median(inputNormalizedDf[[var]], na.rm = T), var])), na.rm = T)
    totalList <-
      c(totalList, 
        exp(nonbinDf[i, 2] * (gMean - lMean)), 
        exp(nonbinDf[i,3] * (gMean - lMean)), 
        exp(nonbinDf[i,4] * (gMean - lMean)))
    
    i <- i + 1
  }
  #create df
  totalList <- unlist(totalList)
  finaldf <- data.frame(0)
  loc <- 1
  for (i in 1:length(listCoef)){
    for (j in 1:3){
      if (j == 1){
        curCol <- paste0("HR_", listCoef[i])
      }
      if (j == 2){
        curCol <- paste0("HR_", listCoef[i], ".LB")
      } 
      if (j == 3){
         curCol <- paste0("HR_", listCoef[i], ".UB")
      }
      finaldf[[curCol]] <- totalList[loc] 
      loc <- loc + 1
    }
  }
  return(select(finaldf, -X0))
}

#FUNCTION for KM plots for given Tumor types.
# PARAMS:
#   km_fit:  survival model fit of data
#   tumorTypeName: name of the tumor type to stratify by
#   coxData:dataset input for the KM graph
KMPlots <- function(km_fit, tumorTypeName, coxData){
  tumor_plot <- ggsurvplot(
    km_fit,
    data = coxData,
    surv.median.line = "hv",
    # Add medians survival
    # Change legends: title & labels
    # legend.title = paste0("T_Type", tumorTypeName, " KM Curves"),
    legend.labs = c(paste0(tumorTypeName, " Negative"), paste0(tumorTypeName, " Positive")),
    ylim = c(0, 1),
    # Add p-value and intervals
    pval = FALSE,
    conf.int = TRUE,
    # Add risk table
    risk.table = TRUE,
    tables.height = 0.2,
    break.time.by = 1000, #to get more granular information
    tables.theme = theme_cleantable(),
    # Color palettes. Use custom color: c("#E7B800", "#2E9FDF"),
    # or brewer color (e.g.: "Dark2"), or ggsci color (e.g.: "jco")
    palette = c("#E7B800", "#2E9FDF"),
    ggtheme = theme_bw() # Change ggplot2 theme
  )
  print(tumor_plot) #output so that we can see out results
  ggsave(plot = print(tumor_plot), paste0(dir, "figures/", "KMPlotTumor", tumorTypeName, ".png"))
}
```


Breast Cancer Dataset (see: https://warwick.ac.uk/fac/sci/statistics/apts/students/resources-1617/gbcs.txt)
```{r}
head(gbcsCS)

# Prepare Data for XGB
BreastCancerDataSet <-as.data.frame(gbcsCS) %>%
  # need to convert to XGB understandable survival times (see XGB documentation for details)
  mutate(SURV_DAY_XGB = ifelse(censdead == 1, survtime, -1*survtime))
# uncomment below to write the resulting file to .csv
write_csv(BreastCancerDataSet, paste0(dir, "data_process/", "XGB_BCDS.csv"))

BreastCancerDataSet_normalized <-BreastCancerDataSet %>% 
  select(id, age, menopause, hormone, size, grade, nodes, prog_recp, estrg_recp, survtime, censdead, SURV_DAY_XGB) %>%
  # normalize the data by min and max values 
  mutate(age = ( (age - min(age)) / (max(age) - min(age))),
         menopause = menopause - 1, 
         hormone = hormone - 1, 
         size = ( (size - min(size)) / (max(size) - min(size))),
         prog_recp = ( (prog_recp - min(prog_recp)) / (max(prog_recp) - min(prog_recp))), 
         estrg_recp = ( (estrg_recp - min(estrg_recp)) / (max(estrg_recp) - min(estrg_recp)))
    )
write_csv(BreastCancerDataSet_normalized, paste0(dir, "data_normalized/", "XGB_BCDS_normalized.csv"))

# normalized CoxPH results
breastCancerCoxPH_normalized <- coxph(Surv(survtime, censdead) ~ ., data = select(BreastCancerDataSet_normalized, -id, -SURV_DAY_XGB))
# save the data
sink(paste0(dir, "results/", "summary_coxPH_normalized_bcd.csv"))
summary(breastCancerCoxPH_normalized)
concordance(breastCancerCoxPH_normalized)
sink()

#Generate materials for the XGB Model comparison
listCoeffients = c("age","menopause","hormone", "size", "grade", "nodes", "prog_recp","estrg_recp")

#full results table (all covs)
results_table_bc <- res_tab(breastCancerCoxPH_normalized, listCoeffients)
write_csv(results_table_bc, paste0(dir, "results/", "summary_coxPH_ALLVars_bcd.csv"))

#Binary covs only
results_binary_bc <- results_table_bc %>% 
  filter(Covariate %in% c('menopause', 'hormone'))
write_csv(results_binary_bc, paste0(dir,"results/", "summary_coxPH_BINARYVARS_bcd.csv"))

#Non-binary covs only
results_nonbin_bc <- results_table_bc %>% 
  filter(Covariate %in% c("age", "size", "grade", "nodes", "prog_recp","estrg_recp")) %>% 
  select(Covariate, Beta, logLB, logUB)
write_csv(results_nonbin_bc, paste0(dir,"results/", "summary_coxPH_NONBINVARS_bcd.csv"))


BCD_nonbin_summary <- nonbin_summary(BreastCancerDataSet_normalized, results_nonbin_bc, list("age", "size", "grade", "nodes", "prog_recp","estrg_recp"))
write_csv(BCD_nonbin_summary, paste0(dir, "results/", "summary_coxPH_nonbinVARS.csv"))


# From: https://stats.stackexchange.com/questions/61090/how-to-split-a-data-set-to-do-10-fold-cross-validation
# Randomly shuffle the data
dataCV <-BreastCancerDataSet_normalized
dataCV <-dataCV[sample(nrow(dataCV)),]

# Create 5 equally size folds
folds <- cut(seq(1,nrow(dataCV)),breaks=5,labels=FALSE)
listCs <- list()
# Perform 5 fold cross validation
for(i in 1:5){
    #Segement your data by fold using the which() function 
    testIndexes <- which(folds==i,arr.ind=TRUE)
    testData <- dataCV[testIndexes, ]
    trainData <- dataCV[-testIndexes, ]
    trained_cox <- (coxph(Surv(survtime, censdead) ~ age + menopause + hormone + size + grade + nodes +  prog_recp + estrg_recp, data = trainData))
    listCs <-
      c(listCs, survConcordance(
        Surv(survtime, censdead) ~ predict(trained_cox, testData),
        testData
      )$concordance)
}
# Print results
for (i in 1:5){
  print((listCs[i]$concordant[1]))
}
#END 5-Fold CV

# For correlation exploration between covariates
BreastCancerDataSetcors <- BreastCancerDataSet %>% 
  select(age, menopause, hormone, size, grade, nodes, prog_recp, estrg_recp)
matrix_cor <- rcorr(as.matrix(BreastCancerDataSetcors), type = "pearson")
# print(matrix_cor)
corrplot(matrix_cor$r, type="upper", order="hclust",method="color", 
         p.mat = matrix_cor$P, sig.level = 0.05, insig = "blank", tl.col="black", tl.srt=45)
  dev.copy(png, paste0(dir, "figures/", "pearsoncorrelogram_bcd.png"))
  dev.off()

  
### To generate same means and sds from figures ###  

# Taken from run of the CoxPH 5-Fold CV
coxCMean <-  mean(c(0.7331452, 0.715131, 0.7260905, 0.7266426, 0.782861))
coxCMean_sd <-  sd(c(0.7331452, 0.715131, 0.7260905, 0.7266426, 0.782861))

# Taken from XGBoost 5-fold CV
xgbCMean <- mean(c(0.7839305103148752,
 0.6783340717766947,
 0.7650669642857143,
 0.7093124456048738,
 0.8002033553634977))
xgbCMean_sd <- sd(c(0.7839305103148752,
 0.6783340717766947,
 0.7650669642857143,
 0.7093124456048738,
 0.8002033553634977))
```




Colon Cancer Dataset (see: https://stat.ethz.ch/R-manual/R-patched/library/survival/html/colon.html)
```{r}
head(colon)
ColonCancerDataSet <-as.data.frame(colon) %>% 
  group_by(id) %>% 
  filter(time == max(time) & etype == 2) %>% 
  mutate(rx = case_when(rx == "Obs" ~ 0, 
                        rx == "Lev" ~ 1, 
                        rx == "Lev+5FU" ~ 2), 
         rx = as.numeric(as.vector(rx)),
    # generate XGB understandable survival times
    SURV_DAY_XGB = ifelse(status == 1, time, -1*time)) %>%
  select(-study, -etype, -node4)


# Normalized Colon Cancer Data
ColonCancerDataSet_normalized <-ColonCancerDataSet %>%
  ungroup() %>% 
  mutate(age = ((age - min(age)) / (max(age) - min(age))),
         nodes = ((nodes - min(nodes, na.rm = T)) / (max(nodes, na.rm = T) - min(nodes, na.rm = T))))
head(ColonCancerDataSet_normalized)
write_csv(ColonCancerDataSet_normalized, paste0(dir, "data_normalized/", "XGB_COLON_normalized.csv"))

# Normalized Colon Cancer CoxPH results
colonCancerCoxPH <- coxph(Surv(time, status) ~ rx + age + sex + obstruct + perfor + adhere + nodes + differ + extent + surg, data = ColonCancerDataSet_normalized)
#save the data
sink(paste0(dir, "results/", "summary_coxPH_normalized_ccd.csv"))
summary(colonCancerCoxPH)
concordance(colonCancerCoxPH)
sink()

# Generate materials for the XGB Model comparison
listCoeffients = c("rx","age", "sex","obstruct",  "perfor",  "adhere",  "nodes" ,"differ",  "extent",  "surg")

# All covs results
results_table_colon <- res_tab(colonCancerCoxPH, listCoeffients)

# Binary only
results_binary_colon <- results_table_colon %>% 
  filter(Covariate %in% c('sex', 'obstruct', 'perfor', 'adhere', 'surg'))
write_csv(results_binary_colon, paste0(dir, "results/", "summary_coxPH_BINARYVARS_ccd.csv"))

# Non-binary only
results_nonbin_colon <- results_table_colon %>% 
  filter(Covariate %in% c("rx","age","nodes" ,"differ",  "extent")) %>% 
  select(Covariate, Beta, logLB, logUB)
write_csv(results_nonbin_colon, paste0(dir, "results/","summary_coxPH_NONBINARYVARS_ccd.csv"))

COLON_nonbin_summary <- nonbin_summary(ColonCancerDataSet_normalized, results_nonbin_colon, list("rx", "age", "nodes" ,"differ", "extent"))
write_csv(COLON_nonbin_summary, paste0(dir,"results/", "summary_coxPH_NONBINVARS_ccd.csv"))

##########
#COLON age cont. test James Nov. 2nd
ColonCancerDataSet_normalized_ageBinary <- ColonCancerDataSet_normalized %>% 
  mutate(age_bin = ifelse(age >= median(age, na.rm=T), 1, 0))
# Normalized Colon Cancer AGE BINARY CoxPH results
colonCancerCoxPH_ageBinary <- coxph(Surv(time, status) ~ rx + age_bin + sex + obstruct + perfor + adhere + nodes + differ + extent + surg, data = ColonCancerDataSet_normalized_ageBinary)
summary(colonCancerCoxPH_ageBinary)
##########

#5-Fold CV 
#From: https://stats.stackexchange.com/questions/61090/how-to-split-a-data-set-to-do-10-fold-cross-validation 
#Randomly shuffle the data
dataCV <-ColonCancerDataSet_normalized
dataCV <-dataCV[sample(nrow(dataCV)),]

#Create 5 equally size folds
folds <- cut(seq(1,nrow(dataCV)),breaks=5,labels=FALSE)
listCs <- list()
#Perform 5 fold cross validation
for(i in 1:5){
    #Segement your data by fold using the which() function 
    testIndexes <- which(folds==i,arr.ind=TRUE)
    testData <- dataCV[testIndexes, ]
    trainData <- dataCV[-testIndexes, ]
    trained_cox <- (coxph(Surv(time, status) ~ rx + age + sex + obstruct + perfor + adhere + nodes + differ + extent + surg, data = trainData))
    listCs <-
      c(listCs, survConcordance(
        Surv(time, status) ~ predict(trained_cox, testData),
        testData
      )$concordance)
    
}
# Print results
for (i in 1:5){
  print((listCs[i]$concordant[1]))
}
#END 5-Fold CV

#Test for correlations
ColonCancerDataSet_cors <- ColonCancerDataSet %>%
  select(-id, -time, -SURV_DAY_XGB)
  
matrix_cor_2 <- rcorr(as.matrix(ColonCancerDataSet_cors), type = "pearson")
# print(matrix_cor)
corrplot(matrix_cor_2$r, type="upper", order="hclust",method="color", 
         p.mat = matrix_cor_2 $P, sig.level = 0.05, insig = "blank", tl.col="black", tl.srt=45)
  dev.copy(png, paste0(dir, "figures/", "pearsoncorrelogram_oct11.png"))
  dev.off()
  
  
### To generate same means and sds from figures ### 

# Taken from run of the CoxPH 5-Fold CV
cVal <- mean(c(0.6098947, 0.6751284, 0.6756624, 0.6192138, 0.6549967))
cVal_std <- sd(c(0.6098947, 0.6751284, 0.6756624, 0.6192138, 0.6549967))

# Taken from run of the XGBoost 5-Fold CV
cValXGB <- mean(c(0.6955977988994497,
 0.6293322628127928,
 0.6540423361387401,
 0.6898670679709055,
 0.6479621290844013))
cValXGB_std <- sd(c(0.6955977988994497,
 0.6293322628127928,
 0.6540423361387401,
 0.6898670679709055,
 0.6479621290844013))
```


Nature Dataset (only Covs used from the Wang thesis, see: https://www.nature.com/articles/nature12634 in supp data 8)
```{r}
#Load data from dataset directory (specified at top)
#Vital.status = 1 if the event of interest is observed, Vital.status = 0 if the observation is censored.
NatureData <- as_tibble(read.csv(paste0(dir, "NatureDataSuppTable.csv"), stringsAsFactors=F), col_types = col_double()) 

NatureData_imputed <- NatureData %>% 
  #convert to binary, impute medians, and exclude patients with no conclusive Survival Data
  mutate(
    Gender.female = ifelse(Gender == "female", 1, 0), 
    Years.to.birth = impute(Years.to.birth, median), 
    Tumor.stage = impute(Tumor.stage, median), 
    # now let's make sure that we have survival times
    time_event = 
      case_when(
        !is.na(Days.to.death) ~ as.numeric(Days.to.death),
        is.na(Days.to.death) &
          !is.na(Days.to.last.followup) ~ as.numeric(Days.to.last.followup),
        T ~ -1
  )) %>%
  filter(time_event > 0) %>% 
  select(-Gender, -Tumor.grade, -Days.to.death, -Days.to.last.followup)
  

# Dataset to add dummy covs for tumor types
NatureData_imputed_dummy <- fastDummies::dummy_cols(NatureData_imputed, 
                                                        select_columns = "TCGA.tumor.type", 
                                                    remove_first_dummy = TRUE)


#From Wang Thesis, "We chose to discard observations with outliers in Somatic.mutations and Nonsilent.somatic.mutations to reduce the noise in the model." Following this, we are aiming to get a dataset that is normalized and close to the results from the Wang thesis.
full_data_nature_normalized <- select(NatureData_imputed_dummy, Sample.ID, Gender.female, Years.to.birth, Vital.status, time_event, Date.of.initial.pathologic.diagnosis, Somatic.mutations, Nonsilent.somatic.mutations, Tumor.stage, TCGA.tumor.type_brca, TCGA.tumor.type_coadread, TCGA.tumor.type_gbm, TCGA.tumor.type_hnsc, TCGA.tumor.type_kirc, TCGA.tumor.type_laml, TCGA.tumor.type_luad, TCGA.tumor.type_lusc, TCGA.tumor.type_ov, TCGA.tumor.type_ucec) %>% 
  # Now, we normalize
  mutate(SURV_DAY_XGB = ifelse(Vital.status == 1, time_event, -1*time_event), 
           Years.to.birth = (Years.to.birth - min(Years.to.birth)) / (max(Years.to.birth) - min(Years.to.birth)), 
         Date.of.initial.pathologic.diagnosis = (Date.of.initial.pathologic.diagnosis - min(Date.of.initial.pathologic.diagnosis)) / (max(Date.of.initial.pathologic.diagnosis) - min(Date.of.initial.pathologic.diagnosis)), 
         Somatic.mutations = (Somatic.mutations - min(Somatic.mutations, na.rm = T)) / (max(Somatic.mutations, na.rm = T) - min(Somatic.mutations, na.rm = T)), 
         Nonsilent.somatic.mutations = (Nonsilent.somatic.mutations - min(Nonsilent.somatic.mutations, na.rm = T)) / (max(Nonsilent.somatic.mutations, na.rm = T) - min(Nonsilent.somatic.mutations, na.rm = T)), 
          Tumor.stage = (Tumor.stage - min(Tumor.stage, na.rm = T)) / (max(Tumor.stage, na.rm = T) - min(Tumor.stage, na.rm = T)))
write_csv(full_data_nature_normalized, paste0(dir, "data_normalized/", "XGB_NATURE_normalized.csv"))
  

# Run the Cox Analysis with the normalized data
NatureCoxPH <- coxph(Surv(time_event, Vital.status) ~ Gender.female + Years.to.birth + Date.of.initial.pathologic.diagnosis + Somatic.mutations + Nonsilent.somatic.mutations + Tumor.stage + TCGA.tumor.type_brca + TCGA.tumor.type_coadread + TCGA.tumor.type_gbm + TCGA.tumor.type_hnsc + TCGA.tumor.type_kirc + TCGA.tumor.type_laml + TCGA.tumor.type_luad + TCGA.tumor.type_lusc + TCGA.tumor.type_ov + TCGA.tumor.type_ucec, data = full_data_nature_normalized)
# Save the data
sink(paste0(dir, "results/", "summary_coxPH_normalized_nature.csv"))
summary(NatureCoxPH)
concordance(NatureCoxPH)
sink()

# Generate materials for the XGB Model comparison
listCoeffients = c("Gender.female", "Years.to.birth", "Date.of.initial.pathologic.diagnosis", "Somatic.mutations", "Nonsilent.somatic.mutations", "Tumor.stage", "TCGA.tumor.type_brca", "TCGA.tumor.type_coadread", "TCGA.tumor.type_gbm", "TCGA.tumor.type_hnsc", "TCGA.tumor.type_kirc", "TCGA.tumor.type_laml", "TCGA.tumor.type_luad", "TCGA.tumor.type_lusc", "TCGA.tumor.type_ov", "TCGA.tumor.type_ucec")

# All Covariates
results_table <- res_tab(NatureCoxPH, listCoeffients)

# Binary covs only
results_binary <- results_table %>% 
  filter(Covariate %in% c("Gender.female", "TCGA.tumor.type_brca", "TCGA.tumor.type_coadread", "TCGA.tumor.type_gbm", "TCGA.tumor.type_hnsc", "TCGA.tumor.type_kirc", "TCGA.tumor.type_laml", "TCGA.tumor.type_luad", "TCGA.tumor.type_lusc", "TCGA.tumor.type_ov", "TCGA.tumor.type_ucec"))
write_csv(results_binary, paste0(dir, "results/", "summary_coxPH_BINARYVARS_natWangOnly.csv"))

# Non-binary covs only 
results_nonbin <- results_table %>% 
  filter(Covariate %in% c("Years.to.birth", "Date.of.initial.pathologic.diagnosis", "Somatic.mutations", "Nonsilent.somatic.mutations", "Tumor.stage")) %>% 
  select(Covariate, Beta, logLB, logUB)

XGB_SUMMARY_TESTFUNCTION <-  nonbin_summary(full_data_nature_normalized,  results_nonbin, list("Years.to.birth", "Date.of.initial.pathologic.diagnosis", "Somatic.mutations", "Nonsilent.somatic.mutations", "Tumor.stage"))
class(XGB_SUMMARY_TESTFUNCTION)
write_csv(XGB_SUMMARY_TESTFUNCTION, paste0(dir, "results/", "summary_coxPH_NONBINVARS_nature_wangonly.csv"))


#5-Fold CV 
#From: https://stats.stackexchange.com/questions/61090/how-to-split-a-data-set-to-do-10-fold-cross-validation 
#Randomly shuffle the data
dataCV <-full_data_nature_normalized
dataCV <-dataCV[sample(nrow(dataCV)),]

#Create 5 equally size folds
folds <- cut(seq(1,nrow(dataCV)),breaks=5,labels=FALSE)
listCs <- list()
#Perform 5 fold cross validation
for(i in 1:5){
    #Segement your data by fold using the which() function 
    testIndexes <- which(folds==i,arr.ind=TRUE)
    testData <- dataCV[testIndexes, ]
    trainData <- dataCV[-testIndexes, ]
    trained_cox <- (coxph(Surv(time_event, Vital.status) ~ Gender.female + Years.to.birth + Date.of.initial.pathologic.diagnosis + Somatic.mutations + Nonsilent.somatic.mutations + Tumor.stage + TCGA.tumor.type_brca + TCGA.tumor.type_coadread + TCGA.tumor.type_gbm + TCGA.tumor.type_hnsc + TCGA.tumor.type_kirc + TCGA.tumor.type_laml + TCGA.tumor.type_luad + TCGA.tumor.type_lusc + TCGA.tumor.type_ov + TCGA.tumor.type_ucec, data = trainData))
    listCs <-
      c(listCs, survConcordance(
        Surv(time_event, Vital.status) ~ predict(trained_cox, testData),
        testData
      )$concordance)
    
}
listCs
#Print the results
for (i in 1:5){
  print((listCs[i]$concordant[1]))
}
#END 5-Fold CV

#Generate KM Curves by tumor types
tumorVar <-  full_data_nature_normalized$TCGA.tumor.type_luad
tumorVar2 <- full_data_nature_normalized$TCGA.tumor.type_lusc
tumorTypeName <- 'LUAD'
tumorTypeName2 <- 'LUSC'
km_fit <- survfit(Surv(time_event, Vital.status) ~ tumorVar, data = full_data_nature_normalized)
km_fit2 <- survfit(Surv(time_event, Vital.status) ~ tumorVar2, data = full_data_nature_normalized)
KMPlots(km_fit, tumorTypeName, full_data_nature_normalized)
KMPlots(km_fit2, tumorTypeName2, full_data_nature_normalized)
```

Nature Data with ALL Covariates 
```{r}
nature_all <- full_data_nature_normalized %>% 
  left_join(select(NatureData_imputed, -Vital.status, -time_event, -TCGA.tumor.type, -Gender.female, -Years.to.birth, -Date.of.initial.pathologic.diagnosis, -Somatic.mutations, -Nonsilent.somatic.mutations, -Tumor.stage), by = "Sample.ID")
write_csv(nature_all, paste0(dir, "data_normalized/", "XGB_NATURE_ALL_normalized.csv"))

coxData <- (select(nature_all, - Sample.ID, -SURV_DAY_XGB))
NatureCoxPH_all <- coxph(Surv(time_event, Vital.status) ~ ., data = coxData)
summary(NatureCoxPH_all)
# Get concordance value from Cox model using "predict"
survConcordance(Surv(time_event, Vital.status) ~ predict(NatureCoxPH_all, coxData), coxData)
#save the data
sink(paste0(dir, "results/", "summary_coxPH_ALL_normalized_natureall.csv"))
summary(NatureCoxPH_all)
concordance(NatureCoxPH_all)
sink()

# Generate materials for the XGB Model comparison 
myColNames <- colnames(select(nature_all, -Sample.ID, -SURV_DAY_XGB, -time_event, -Vital.status))

# All covs
results_table_all <- res_tab(NatureCoxPH_all, myColNames)
write_csv(results_table_all, paste0(dir, "results/","summary_coxPH_ALL_natureall.csv"))
# Non-binary covs
results_nonbin_all <- results_table_all %>% 
  filter(Covariate %in% c("Years.to.birth", "Date.of.initial.pathologic.diagnosis", "Somatic.mutations", "Nonsilent.somatic.mutations", "Tumor.stage")) %>% 
  select(Covariate, Beta, logLB, logUB)
write_csv(results_nonbin_all, paste0(dir, "results/","summary_coxPH_NONBINVARS_ALL_natureall.csv"))

# Binary covs
results_binary_all <- results_table_all %>% 
  anti_join(results_nonbin_all, by = "Covariate") %>% 
  select(Covariate, HR, LB, UB)
write_csv(results_binary_all, paste0(dir,"results/", "summary_coxPH_BINARYVARS_ALL_natureall.csv"))

XGB_SUMMARY_ALL <- nonbin_summary(nature_all, results_nonbin_all, list("Years.to.birth", "Date.of.initial.pathologic.diagnosis", "Somatic.mutations", "Nonsilent.somatic.mutations", "Tumor.stage"))
write_csv(XGB_SUMMARY_ALL, paste0(dir,"results/", "summary_coxPH_NONBINVARS_ALL_natureall.csv"))

#5-Fold CV
#From: https://stats.stackexchange.com/questions/61090/how-to-split-a-data-set-to-do-10-fold-cross-validation 
#Randomly shuffle the data
dataCV <-coxData
dataCV <-dataCV[sample(nrow(dataCV)),]

#Create 5 equally size folds
folds <- cut(seq(1,nrow(dataCV)),breaks=5,labels=FALSE)
listCs <- list()
#Perform 5 fold cross validation
for(i in 1:5){
    #Segement your data by fold using the which() function 
    testIndexes <- which(folds==i,arr.ind=TRUE)
    testData <- dataCV[testIndexes, ]
    trainData <- dataCV[-testIndexes, ]
    trained_cox <- (coxph(Surv(time_event, Vital.status) ~ ., data = trainData))
    listCs <-
      c(listCs, survConcordance(
        Surv(time_event, Vital.status) ~ predict(trained_cox, testData),
        testData
      )$concordance)
    
}
listCs
#Print values
for (i in 1:5){
  print((listCs[i]$concordant[1]))
}


tumorVar3 <- coxData$TCGA.tumor.type_gbm
km_fit3 <- survfit(Surv(time_event, Vital.status) ~ tumorVar3, data = coxData)
KMPlots(km_fit3, 'GBM', coxData)

#Generate KM Curves by tumor types
tumorVar <-  coxData$TCGA.tumor.type_luad
tumorVar2 <- coxData$TCGA.tumor.type_lusc
tumorTypeName <- 'LUAD_ALL'
tumorTypeName2 <- 'LUSC_ALL'
km_fit <- survfit(Surv(time_event, Vital.status) ~ tumorVar, data = coxData)

KMPlots(km_fit, tumorTypeName, coxData)
KMPlots(km_fit2, tumorTypeName2, coxData)

### To generate same means and sds from figures ### 

#Get average from results
mean(c(0.7466364, 0.7911606, 0.7679637, 0.7341046, 0.7471551))

```


Nature, all data, no imputations for XGB to see comparative performance
```{r}
nature_all_normalized_noimputations <- NatureData %>%
  mutate(
    Gender.female = ifelse(Gender == "female", 1, 0),
    # now let's make sure that we have survival times
    time_event =
      case_when(
        !is.na(Days.to.death) ~ as.numeric(Days.to.death),
        is.na(Days.to.death) &
          !is.na(Days.to.last.followup) ~ as.numeric(Days.to.last.followup),
        T ~ -1
      ),
    SURV_DAY_XGB = ifelse(Vital.status == 1, time_event, -1 * time_event),
    Years.to.birth = (Years.to.birth - min(Years.to.birth, na.rm = T)) / (max(Years.to.birth, na.rm = T) - min(Years.to.birth, na.rm = T)),
    Date.of.initial.pathologic.diagnosis = (
      Date.of.initial.pathologic.diagnosis - min(Date.of.initial.pathologic.diagnosis, na.rm = T)
    ) / (
      max(Date.of.initial.pathologic.diagnosis, na.rm = T) - min(Date.of.initial.pathologic.diagnosis, na.rm = T)
    ),
    Somatic.mutations = (Somatic.mutations - min(Somatic.mutations, na.rm = T)) / (
      max(Somatic.mutations, na.rm = T) - min(Somatic.mutations, na.rm = T)
    ),
    Nonsilent.somatic.mutations = (
      Nonsilent.somatic.mutations - min(Nonsilent.somatic.mutations, na.rm = T)
    ) / (
      max(Nonsilent.somatic.mutations, na.rm = T) - min(Nonsilent.somatic.mutations, na.rm = T)
    ),
    Tumor.stage = (Tumor.stage - min(Tumor.stage, na.rm = T)) / (max(Tumor.stage, na.rm = T) - min(Tumor.stage, na.rm = T)),
  ) %>%
  filter(time_event > 0) %>%
  select(-Gender,-Tumor.grade,-Days.to.death,-Days.to.last.followup) %>%
  fastDummies::dummy_cols(select_columns = "TCGA.tumor.type", remove_first_dummy = TRUE) %>% 
  select(-TCGA.tumor.type)

write_csv(nature_all_normalized_noimputations, paste0(dir, "data_normalized/", "XGB_NATURE_ALL_noimputing_normalized.csv"))

```

Generate frequency tables for significant XGB and COX Covariates
```{r}
SigCols <- c(
 'TCGA.tumor.type_brca',
 'TCGA.tumor.type_coadread',
 'TCGA.tumor.type_gbm',
 'TCGA.tumor.type_hnsc',
 'TCGA.tumor.type_kirc',
 'TCGA.tumor.type_laml',
 'TCGA.tumor.type_luad',
 'TCGA.tumor.type_lusc',
 'TCGA.tumor.type_ov',
 'TCGA.tumor.type_ucec',
 'AR',
 'BAP1',
 'CDKN1A',
 'CDKN1B',
 'DNMT3A',
 'EGFR',
 'ERCC2',
 'EZH2',
 'FBXW7',
 'HIST1H1C',
 'IDH1',
 'PPP2R1A',
 'RAD21',
 'RPL22',
 'RUNX1',
 'TP53')

freqtables <- nature_all %>% 
  select(SigCols)
colSums(is.na(freqtables))

freq_tables_final <- apply(freqtables, 2, table)
write_csv(as.data.frame(freq_tables_final), paste0(dir, "freqTable.csv"))
```

```{r}
### To generate same means and sds from figures ### 

#Get cIndex Vals 
Cox_all <-mean(c(0.7466364, 0.7911606, 0.7679637, 0.7341046, 0.7471551))
Cox_limitedcovs <- mean(c(0.7968867, 0.7813753, 0.7727417, 0.7902709, 0.7645715))
Cox_all_std <-sd(c(0.7466364, 0.7911606, 0.7679637, 0.7341046, 0.7471551))
Cox_limitedcovs_std <- sd(c(0.7968867, 0.7813753, 0.7727417, 0.7902709, 0.7645715))
xgb_lim_mean <- mean(c(0.7832576249188838,
 0.7938660682342993,
 0.7590855642757527,
 0.7545489972721005,
 0.7585598214053941))
xgb_all_mean <- mean(c(0.7820116807268008,
 0.7986470224737373,
 0.7598238945554978,
 0.7561977277496328,
 0.7571580614178542))
xgb_nImp_mean <- mean(c(0.7800129785853342,
 0.7998829154063811,
 0.766468867073204,
 0.7653107107527204,
 0.7605326687952652))
xgb_lim_std <- sd(c(0.7832576249188838,
 0.7938660682342993,
 0.7590855642757527,
 0.7545489972721005,
 0.7585598214053941))
xgb_all_std <- sd(c(0.7820116807268008,
 0.7986470224737373,
 0.7598238945554978,
 0.7561977277496328,
 0.7571580614178542))
xgb_nImp_std <- sd(c(0.7800129785853342,
 0.7998829154063811,
 0.766468867073204,
 0.7653107107527204,
 0.7605326687952652))
```


